---
title: 'Final project'
output: github_document
---

```{r}
library(tidyverse)
library(GGally)
library(patchwork)
library(gt)
library(leaps)
library(caret)
library(readxl)
library(patchwork)
library(glmnet)
library(performance)
```

```{r}
body_density_df = read_excel("data/body_density_data.xlsx") %>%
  rename(outcome = body_density) %>% 
  select(-bodyfat_siri & -bodyfat_brozek)
head(body_density_df)
```

Descriptive statistics:

```{r}
body_density_df %>%
  gtsummary::tbl_summary() %>%
  gtsummary::bold_labels()

summary(body_density_df)
```

Exploratory plots

```{r}
correlation_plot = body_density_df %>%
  relocate(outcome) %>%
  ggpairs()
  ggsave("./correlation.png",correlation_plot, width = 8, height = 5)
```

Histogram plots

```{r}
hist1 = body_density_df %>% 
  ggplot(aes(x = age)) +
  geom_histogram()
hist2 = body_density_df %>% 
  ggplot(aes(x = weight)) +
  geom_histogram()
hist3 = body_density_df %>% 
  ggplot(aes(x = height)) +
  geom_histogram()

hist1 + hist2 + hist3

hist4 = body_density_df %>% 
  ggplot(aes(x = neck)) +
  geom_histogram()
hist5 = body_density_df %>% 
  ggplot(aes(x = chest)) +
  geom_histogram()
hist6 = body_density_df %>% 
  ggplot(aes(x = abdomen)) +
  geom_histogram()

hist4 + hist5 + hist6

hist7 = body_density_df %>% 
  ggplot(aes(x = hip)) +
  geom_histogram()
hist8 = body_density_df %>% 
  ggplot(aes(x = thigh)) +
  geom_histogram()
hist9 = body_density_df %>% 
  ggplot(aes(x = knee)) +
  geom_histogram()

hist7 + hist8 + hist9

hist10 = body_density_df %>% 
  ggplot(aes(x = ankle)) +
  geom_histogram()
hist11 = body_density_df %>% 
  ggplot(aes(x = bicep)) +
  geom_histogram()
hist12 = body_density_df %>% 
  ggplot(aes(x = forearm)) +
  geom_histogram()

hist10 + hist11 + hist12

hist13 = body_density_df %>% 
  ggplot(aes(x = wrist)) +
  geom_histogram()

hist13
```

## Fitting a full model and checking diagnostic plots

```{r}
full_model = lm(outcome ~ ., data = body_density_df)
summary(full_model)

par(mfrow = c(2,2))
plot(full_model)
```

The diagnostics plots look good. No need for Box-Cox transformation

## Forward selection

```{r}
intercept_only <- lm(outcome ~ 1, data = body_density_df)
fit_forward = step(intercept_only, direction = "forward", scope = formula(full_model))
```

## Backward Elimination

```{r}
fit_backward = step(lm(full_model, data = body_density_df), direction = "backward")
```

## Stepwise selection

```{r}
step(
lm(outcome ~ age + weight + height + neck + chest + abdomen + hip + knee + ankle + bicep + forearm + wrist,
data = body_density_df),
direction = "both"
) 
```

## Criteria-based Selection

### Plot

```{r}
b = regsubsets(outcome ~ 
                 age + weight + height + neck + chest + abdomen + hip + thigh +
                 knee + ankle + bicep + forearm + wrist, data = body_density_df)
rs = summary(b)
rs$which

par(mfrow=c(1,2))
plot(2:9, rs$cp, xlab = "No of parameters", ylab = "Cp Statistic")
abline(0,1)
   
plot(2:9, rs$adjr2, xlab = "No of parameters", ylab = "Adj R2")
par(mfrow = c(1,1))
```

### cp and adjr2 values

 age + weight + height + neck + chest + abdomen + hip + thigh +
                 knee + ankle + bicep + forearm + wrist
```{r}
cp_model =  leaps(x = as.matrix(body_density_df)[,c(3:15)], y = as.matrix(body_density_df)[,2], nbest = 1, method = "Cp")
cp_model$which
3591011

ggplot(data.frame(cp_model$size, cp_model$Cp), aes(x = cp_model.size, y = cp_model.Cp)) +  labs(y = "Cp", x = "# parameters") + scale_x_continuous(breaks=seq(1,14)) + geom_point()

adjr2_model =  leaps(x = as.matrix(body_density_df)[,c(3:15)], y = as.matrix(body_density_df)[,2], nbest = 1, method = "adjr2")
adjr2_model$which
359
```

Best model
```{r}
fit_cp = lm(outcome ~ age + weight + neck + abdomen + hip + thigh 
                    + forearm + wrist, data = body_density_df)
fit_cp

fit_adjr2 = lm(outcome ~ age + weight + neck + abdomen + hip + thigh + ankle + bicep + forearm + wrist, data = body_density_df)
fit_adjr2
```


## Lasso

```{r}
new_df = body_density_df %>% 
  select(-id)
# using cross validation to choose lambda
lambda_seq <- 10^seq(-3, 0, by = .1)
set.seed(2022)
cv_object <- glmnet::cv.glmnet(as.matrix(new_df[2:14]), new_df$outcome, 
                       lambda = lambda_seq, 
                       nfolds = 5)
cv_object 

# plot the CV results
tibble(lambda = cv_object$lambda,
       mean_cv_error = cv_object$cvm) %>%
  ggplot(aes(x = lambda, y = mean_cv_error)) +
  geom_point()

# extracting the exact minimum lambda from the CV object
cv_object$lambda.min


# refit the lasso model with the "best" lambda
fit_bestcv <- glmnet::glmnet(as.matrix(new_df[2:14]), new_df$outcome, lambda = cv_object$lambda.min)
coef(fit_bestcv)
```

```{r}
set.seed(1)
# Use 10-fold validation and create the training sets
train = trainControl(method = "cv", number = 10)

# Fit the 4-variables model that we discussed in previous lectures
forward_caret = train(outcome ~ abdomen + weight + wrist + forearm + bicep + neck,
                  data = body_density_df,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)

forward_caret$finalModel
print(forward_caret)
```

```{r}
# Use 10-fold validation and create the training sets
train = trainControl(method = "cv", number = 10)

# Fit the 4-variables model that we discussed in previous lectures
backward_caret = train(outcome ~ age + weight + neck + abdomen + hip + thigh + bicep + forearm + wrist,
                  data = body_density_df,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)

backward_caret$finalModel
print(backward_caret)
```

```{r}
# Use 10-fold validation and create the training sets
train = trainControl(method = "cv", number = 10)

# Fit the 4-variables model that we discussed in previous lectures
stepwise_caret = train(outcome ~ weight + neck + abdomen + bicep + forearm + 
    wrist,
                  data = body_density_df,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)

stepwise_caret$finalModel
print(stepwise_caret)
```

```{r}
# Use 10-fold validation and create the training sets
train = trainControl(method = "cv", number = 10)

# Fit the 4-variables model that we discussed in previous lectures
cp_caret = train(outcome ~ age + weight + neck + abdomen + hip + thigh + forearm + wrist,
                  data = body_density_df,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)

cp_caret$finalModel
print(cp_caret)
```


```{r}
# Use 10-fold validation and create the training sets
train = trainControl(method = "cv", number = 10)

# Fit the 4-variables model that we discussed in previous lectures
adjr2_caret = train(outcome ~ age + weight + neck + abdomen + hip + thigh + ankle + bicep + forearm + wrist,
                  data = body_density_df,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)

adjr2_caret$finalModel
print(adjr2_caret)
```

```{r}
# Use 10-fold validation and create the training sets
lasso = trainControl(method = "cv", number = 10)

# Fit the 4-variables model that we discussed in previous lectures
lasso_caret = train(outcome ~ age + height + abdomen + wrist,
                  data = body_density_df,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)

lasso_caret$finalModel
print(lasso_caret)
```

```{r}
best_model=lm(outcome ~ weight + neck + abdomen + bicep + forearm + wrist, data = body_density_df)
```


```{r}
plot(best_model, which = 4, id.n = 3)
```


